Toward \earlywarning\ detection of gravitational waves from compact binary coalescence

### 2012-01-05 article.tex
Added square to numerator in definition of signal to noise
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1440 ccdd05d5-56c1-461a-bf98-85f6876523b4
Rearranged clauses in sentence
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1441 ccdd05d5-56c1-461a-bf98-85f6876523b4

Searches for inspiral signals typically employ matched filter
banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.  Suppose that the observed data $x[k]$ consists of a known, nominal signal $s[k]$ and additive, zero-mean noise $n[k]$,
%
$$
	x[k] = s[k] + n[k].
$$
%
A matched filter is a linear filter, defined as
%
$$
	y[k] = \sum_{n=0}^{N-1} h[n] \, x[k-n] = y_s[k] + y_n[k],
$$
%
where $y_s$ is the response of the filter to the signal alone and $y_n$ is the response of the signal to noise alone.  The matched filter's coefficients maximize the ratio of the expectation of the filter's instantaneous response to the variance in the filter's output:
%
$$
\textrm{signal to noise} = \frac{\expectation \left[ y[0] \right]}{\var \left[ y[k] \right]} = \frac{y_s[0]}{\var \left[ y_n[k] \right]}.
$$
%
It is well known~\citep[see, for example,][]{matched-filter} that if $n[k]$ is Gaussian and wide-sense stationary, then the optimum is obtained when
%
$$
\tilde{h}[n] = \tilde{s}^*[n] \, \tilde{S}^{-1}[n],
$$
%
up to an arbitrary multiplicative constant.  Here, $\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms (DFTs) of $h[k]$, $s[k]$, and $x[k]$, respectively; $\tilde{S}[n] = \expectation \left[ \tilde{n}[n] \tilde{n}^* [n] \right]$ is the folded, two-sided, discrete power spectrum of $n[k]$.  It is related to the continuous, one-sided power spectral density $S(f)$ through
$$
	\tilde{S}[n] =
	\begin{cases}
		S(n f_s / 2 N) / 2 & \textrm{if } 0 \leqslant n \leqslant N / 2 \\
		\tilde{S}[N - n] & \textrm{otherwise},
	\end{cases}
$$
where $N$ is the length of the filter and $f_s$ is the sample rate.  (In order to satisfy the Nyquist-Shannon sampling criterion, it is assumed that the detector's continuous noise power spectral density $S(f)$ vanishes for all $f > f_s / 2$, or alternatively, that the data is low-pass filtered prior to matched filtering.)  The discrete Fourier transform of the output is
%
\begin{multline}
\label{eq:matched-filter-fd}
\tilde{y}[n] = \tilde{s}^*[n] \, \tilde{S}^{-1}[n] \, \tilde{x}[n] \\
\equiv \left(\tilde{S}^{-1/2}[n] \, \tilde{s}[n]\right)^* \left(\tilde{S}^{-1/2}[n] \, \tilde{x}[n] \right).
\end{multline}
%
The placement of parentheses in equation~(\ref{eq:matched-filter-fd}) emphasizes that the matched filter can be thought of as a cross-correlation of a whitened version of the data with a whitened version of the nominal signal.  In this paper, we shall not describe the exact process by which the detector's noise power spectrum is estimated and deconvolved from the data; for the remainder of this paper we shall define $x[k]$ as the \emph{whitened} data stream.  Correspondingly, from this point on we shall use $h[k]$ to describe the \emph{whitened} templates, being the inverse DFT of $\left(\tilde{S}^{-1/2}[n] \, \tilde{s}[n]\right)^*$.

@@

The first step of our proposed method is to divide the templates into time
slices in a \TD\ analogue to the \FD\ decomposition employed by MBTA~\citep{Marion2004,Buskulic2010}.  The application to GW data analysis is foreshadowed by an earlier \FD\ convolution algorithm based on splitting the impulse response of a filter into smaller blocks, proposed by \citet{gardner1995efficient}.  We decompose each template
$h_{i}[k]$ into a sum of $S$ non-overlapping templates
%
\begin{equation}
\label{eq:time-slices}
h_{i}[k] = \sum_{s=0}^{S-1}
	\begin{cases}
		h_i^s[k] & \textrm{if } t^s \leqslant k / f^0 < t^{s+1} \\
		0 & \textrm{otherwise}
	\end{cases}
\end{equation}
%
for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0
t^S = N$.  The outputs of these new time-sliced filters
form an ensemble of partial \SNR\ streams.  By linearity of the filtering
process, these partial \SNR\ streams can be summed to reproduce the
\SNR\ of the full template.

>>>>

Searches for inspiral signals typically employ matched filter
banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.  Suppose that the observed data $x[k]$ consists of a known, nominal signal $s[k]$ and additive, zero-mean noise $n[k]$,
%
$$
	x[k] = s[k] + n[k].
$$
%
A matched filter is a linear filter, defined as
%
$$
	y[k] = \sum_{n=0}^{N-1} h[n] \, x[k-n] = y_s[k] + y_n[k],
$$
%
where $y_s$ is the response of the filter to the signal alone and $y_n$ is the response of the signal to noise alone.  The matched filter's coefficients maximize the ratio of the expectation of the filter's instantaneous response to the variance in the filter's output:
%
$$
\textrm{signal to noise} = \frac{\expectation \left[ y[0] \right]^2}{\var \left[ y[k] \right]} = \frac{y_s[0]^2}{\var \left[ y_n[k] \right]}.
$$
%
It is well known~\citep[see, for example,][]{matched-filter} that if $n[k]$ is Gaussian and wide-sense stationary, then the optimum is obtained when
%
$$
\tilde{h}[n] = \tilde{s}^*[n] \, \tilde{S}^{-1}[n],
$$
%
up to an arbitrary multiplicative constant.  Here, $\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms (DFTs) of $h[k]$, $s[k]$, and $x[k]$, respectively; $\tilde{S}[n] = \expectation \left[ \tilde{n}[n] \tilde{n}^* [n] \right]$ is the folded, two-sided, discrete power spectrum of $n[k]$.  It is related to the continuous, one-sided power spectral density $S(f)$ through
$$
	\tilde{S}[n] =
	\begin{cases}
		S(n f_s / 2 N) / 2 & \textrm{if } 0 \leqslant n \leqslant N / 2 \\
		\tilde{S}[N - n] & \textrm{otherwise},
	\end{cases}
$$
where $N$ is the length of the filter and $f_s$ is the sample rate.  (In order to satisfy the Nyquist-Shannon sampling criterion, it is assumed that the detector's continuous noise power spectral density $S(f)$ vanishes for all $f > f_s / 2$, or alternatively, that the data is low-pass filtered prior to matched filtering.)  The discrete Fourier transform of the output is
%
\begin{multline}
\label{eq:matched-filter-fd}
\tilde{y}[n] = \tilde{s}^*[n] \, \tilde{S}^{-1}[n] \, \tilde{x}[n] \\
\equiv \left(\tilde{S}^{-1/2}[n] \, \tilde{s}[n]\right)^* \left(\tilde{S}^{-1/2}[n] \, \tilde{x}[n] \right).
\end{multline}
%
The placement of parentheses in equation~(\ref{eq:matched-filter-fd}) emphasizes that the matched filter can be thought of as a cross-correlation of a whitened version of the data with a whitened version of the nominal signal.  In this paper, we shall not describe the exact process by which the detector's noise power spectrum is estimated and deconvolved from the data; for the remainder of this paper we shall define $x[k]$ as the \emph{whitened} data stream.  Correspondingly, from this point on we shall use $h[k]$ to describe the \emph{whitened} templates, being the inverse DFT of $\left(\tilde{S}^{-1/2}[n] \, \tilde{s}[n]\right)^*$.

@@

The first step of our proposed method is to divide the templates into time
slices in a \TD\ analogue to the \FD\ decomposition employed by MBTA~\citep{Marion2004,Buskulic2010}.  The application to GW data analysis is foreshadowed by an earlier \FD\ convolution algorithm, proposed by \citet{gardner1995efficient}, based on splitting the impulse response of a filter into smaller blocks.  We decompose each template
$h_{i}[k]$ into a sum of $S$ non-overlapping templates
%
\begin{equation}
\label{eq:time-slices}
h_{i}[k] = \sum_{s=0}^{S-1}
	\begin{cases}
		h_i^s[k] & \textrm{if } t^s \leqslant k / f^0 < t^{s+1} \\
		0 & \textrm{otherwise}
	\end{cases}
\end{equation}
%
for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0
t^S = N$.  The outputs of these new time-sliced filters
form an ensemble of partial \SNR\ streams.  By linearity of the filtering
process, these partial \SNR\ streams can be summed to reproduce the
\SNR\ of the full template.
